Clustering in a hyperbolic model of complex networks.

### 2019-05-01 transition.tex
processed most of Nikoloas comments for sections 1-3
processed most of Nikolaos comments on sections 5-7

Recall that $\Gbox$ is obtained by restricting the Poisson Point Process $\PPP$ to the box $\Rcal = (-I_n, I_n] \times (0, R_n]$, with $I_n = \frac{\pi}{2} e^{R_n/2}$ and connecting two points $p_1, p_2 \in \Rcal$ if $|x_1 - x_2|_{\pi e^{R_n/2}} \le e^{(y_1 + y_2)/2}$. We also recall that by definition of the norm $|.|_{\pi e^{R_n/2}}$ the left and right boundaries of $\Rcal$ are identified. See Section~\ref{ssec:finite_model} for more details. Due to this identification of the boundaries some triples of nodes that form triangles in the finite box model do not form a triangle in the infinite model. Therefore, to establish the required result we need to compute the asymptotic difference between triangle counts in both models.

@@

%We now turn to the task of calculating the expected number of triangles of a node at height $y$, for both the infinite model and the finite box model. 
For any $p \in \R \times \R_+$ we define for the finite box model,
\[
	T_{\text{box}}(p) = \sum^{\ne}_{p_1, p_2 \in \Pcal_n \setminus p} T_{\text{box}}(p,p_1,p_2)
\]
where the sum is over all distinct pairs in $\Pcal_n \setminus p$ and
\[
	T_{\text{box}}(p,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPon{p_1}}.
\]
Similarly, for the infinite model we define
\[
	T_\infty(y) = \sum_{p_1, p_2 \in \Pcal \setminus \{0,y\}}^{\ne} T_\infty(y,p_1,p_2),
\]
where
\[
	T_{\infty}(y,p_1,p_2) = \ind{p_1 \in \BallPo{y}}\ind{p_2 \in \BallPo{y}}\ind{p_2 \in \BallPo{p_1}}.
\]
We will first relate $\gamma(k_n)$ and $\Exp{c^\ast(k_n; \Gbox)}$ using $T_\infty(y)$ and $T_{\text{box}}(y)$. Recall the definition of $\Kcal_{C}(k_n)$
\[
	\Kcal_C(k_n) = \left\{y \in \R_+ : \frac{k_n - C \sqrt{k_n \log(k_n)}}{\xi} \le e^{\frac{y}{2}}
	\le \frac{k_n + C \sqrt{k_n \log(k_n)}}{\xi} \right\},	
\]

@@

For~\eqref{eq:alt_c_ast_kn} we recall that
\[
	c^\ast(k_n; \Gbox) = \frac{1}{\Exp{\Nbox(k_n)}} \sum_{p \in \Pcal} c_{\text{box}}(p)\ind{\Dbox(p) = k_n},
\]
where $c_{\text{box}}(p)$ can be expressed as
\[
	c_{\text{box}}(p) = \frac{1}{\binom{\Dbox(p)}{2}} \sum_{p_1, p_2 \in \Pcal \setminus p}^{\ne} T_{\text{box}}(p,p_1,p_2)
	= \frac{T_{\text{box}}(p)}{\binom{\Dbox(p)}{2}}.
\]
By the Campbell-Mecje formula
\begin{align*}
	\Exp{c^\ast(k_n; \Gbox)} 
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \Exp{c_{\text{box}}(p)\ind{\Dbox(p) = k_n}} f(x,y) \dd x \dd y\\
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \CExp{c_{\text{box}}(p)}{\Dbox(p) = k_n}\rho_{\text{box}}(p,k_n) 
		f(x,y) \dd x \dd y\\
	&= (1+\smallO{1})\frac{n}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
		\rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line follows from a concentration of heights argument, for which we used that $\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n} \le 1$. To analyze the conditional expectation we observe that, similar to the analysis of $\gamma(k_n)$, conditioned on their being $k_n$ points in $\BallPon{y}$, each points $u_i = (x_i,y_i)$ is independently distributed according to the probability density $\Mu{\BallPon{y}}^{-1} \ind{u_i \in \BallPon{y}} f(x_i,y_i)$. Therefore,
\begin{align*}
	\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
	&= \binom{k_n}{2}^{-1} \Exp{\sum_{1 \le i < j \le k_n} \ind{u_i \in \BallPon{u_j}}} \\
	&= \Exp{u_1 \in \BallPon{u_2}} \\
	&= \Mu{\BallPon{y}}^{-2} \iint T_{\text{box}}(y,p_1,p_2) f(x_1,y_1) f(x_2,y_2) 
		\dd x_1 \dd y_1 \dd x_2 \dd y_2\\
	&= \Mu{\BallPon{y}}^{-2} \Exp{T_{\text{box}}(y)}.
\end{align*}
and thus, by applying a concentration of heights argument on $\Mu{\BallPon{y}}^{-2}$,
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{n \Mu{\BallPon{2\log(k_n/\xi)}}^{-2}}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \Exp{T_{\text{box}}(y)} \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
To finish the argument, we first note that $\Mu{\BallPon{2\log(k_n/\xi)}}^{-2} = (1+\smallO{1})k_n^2$, while
\[
	\Exp{N_{\text{box}}(k_n)} = \int_{\Rcal} \rho_{\text{box}}(y,k_n) f(x,y) \dd x \dd y,
\]
so that by a concentration of heights argument,
\[
	\Exp{N_{\text{box}}(k_n)} = (1+o(1)) n \int_{0}^\infty \rho(y,k_n) \alpha e^{-\alpha y} \dd y 
	= (1+\smallO{1}) n p_{k_n}.
\]
We therefore conclude that
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{1}{k_n^2 p_{k_n}} \int_{\Kcal_C(k_n)} 
		\Exp{T_{\text{box}}(y)} \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\]

@@

To analyze $T_{\text{box}}(y_0) - T_\infty(y_0)$ we first reiterate that the difference between the indicator $\ind{p_1 \in \BallPon{p}}$ in the finite box model and $\ind{p_1 \in \BallPo{p}}$ is that in $\Gbox$ we identified the boundaries of the interval $[-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2} e^{R_n/2}]$ and we stop at height $y = R_n$. This induces a difference in triangle counts between both models. 
To see this, note that for any $p = (x,y)$ with $0 \le y \le R_n$ we have that $\BallPon{p} = \BallPo{p} \cap \Rcal$. This means that if $p^\prime, p_2 \in \BallPon{p}$ and $p_2 \in \BallPo{p^\prime} \cap \Rcal$ then $p_2 \in \BallPon{p} \cap \BallPon{p^\prime}$ and hence $(p,p^\prime,p_2)$ form a triangle both in $\Gbox$ and $\Ginf$. However, it could happen that there are points in the intersection $\BallPon{p} \cap \BallPon{p^\prime}$ that are not in $\BallPo{p} \cap \BallPo{p^\prime}$. Let us denote this region by $\mathcal{T}(p \Delta p^\prime)$, see Figure~\ref{fig:comparing_triangles_diff_intersections} for an example of this region. Then, any $p_2 \in \mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$ creates a triangle with $p$ and $p^\prime$ in $\Gbox$ that is not present in $\Ginf$. Finally, any point $p_2 \in \BallPo{p} \cap \BallPo{p6\prime}$ with height $y_2 > R$ creates a triangle with $p, p^\prime$ in $\Ginf$ but not in $\Gbox$.

Let us now define the following triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus p_0}^{\ne} 
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal}.
\]
Then $\widetilde{T}_{\text{box}}(p_0)$ only counts those triangles attached to $p_0$ that exist in both $\Gbox$ and $\Ginf$ and thus, by definition of the region $\mathcal{T}(p_0,p_1)$,
\[
	T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)
	= \sum_{p_1, p_2 \in \Pcal }^{\ne} \ind{p_1 \in \BallPon{p_0}} 
			\ind{p_2 \in \mathcal{T}(p_0, p_1)}.
\]
%Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.
%Observe that the difference between $\widetilde{T}_{\text{box}}(p_0)$ and $T_{\text{box}}(p_0)$ is in the last indicator in $\widetilde{T}_{\text{box}}(p_0,p_1,p_2)$.  counts those points $p_2$ in the box $\Rcal$ that are in the neighborhood $\BallPo{p_1}$ of $p_1$ in the infinite model. Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.

The next result, which is crucial for the proof of Proposition~\ref{prop:convergence_average_clustering_P_n}, computes the expected measure of $\mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$ with respect to $p^\prime$.

\begin{lemma}\label{lem:clustering_error_T_term}
Let $p_0 = (0,y)$ with $y \in \Kcal_{C}(k_n)$. Then as $n \to \infty$,
\[
	\Exp{\left|T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right|}
	= y \, \bigO{n^{-(2\alpha - 1)}} + e^{y}\,\bigO{n^{-(4\alpha-2)}}.
\]
%\[
%	\int_{\Rcal} \mu\left(\mathcal{T}_{\Pcal \Delta \Pcal_n}(p_0,p_1)\right) f(x_1,y_1) 
%	\dd x_1 \dd y_1 = \bigO{y n^{-(2\alpha - 1)} + n^{-(2\alpha-1)} e^{y}}
%\]
\end{lemma}

@@

\begin{proof}[Proof of Proposition~\ref{prop:convergence_average_clustering_P_n}]
Let us write $\Rcal^\prime := \R \times \R_+ \setminus \Rcal$ and let $p_0 = (0,y)$ denote the typical point. Next we recall that it is enough to show~\eqref{eq:triangle_diff_main_term}, so that in particular we have that $y \in \Kcal_{C}(k_n)$.

@@

\TM{ in the 1st integral we could have restricted to the ball of $p$, and possibly saved a lot. Does that not help? }\PvdH{the first integral is taken care of by another Lemma so it does not matter I think.}
The part is taken care of by Lemma~\ref{lem:clustering_error_T_term}. For the other integral we have
\begin{align*}
	&\hspace{-30pt}\iint_{\Rcal^\prime} T_{\infty}(p_0,p_1,p_2) f(x_1,y_1) f(x_2,y_2)
		\dd x_2 \dd y_2 \dd x_1 \dd y_1\\
	&\le \left(\int_{\Rcal^\prime} \ind{p_1 \in \BallPo{p_0}} f(x_1,y_1) \dd x_1 \dd y_1\right)^2
		= \bigO{\left(e^{y/2} \int_{R_n}^\infty e^{-(\alpha - \frac{1}{2})y_1} \dd y_1\right)^2}\\
	&= \bigO{e^{y} e^{-(2\alpha-1)R_n}} = \bigO{e^y n^{-(4\alpha - 2)}}.
\end{align*}
%\TM{ Typo : $f_{\mu, \nu}$? Also give more details how line 3 follows. }
We thus we conclude, using Lemma~\ref{lem:clustering_error_T_term}, that,
\begin{equation}\label{eq:triangle_count_diff_scaling}
	\left|\Exp{T_{\text{box}}(p_0) - T_{\infty}(p_0)}\right| = \bigO{y n^{-(2\alpha - 1)} + n^{-(4\alpha-2)} e^{y}}.
\end{equation}
Therefore, since $y = \bigO{\log(k_n)}$ on $\Kcal_C(k_n)$, 
\begin{align*}
	&\hspace{-30pt}\int_{\Kcal_{C}(k_n)} \rho(y,k_n) \left|\Exp{T_{\text{box}}(p_0) - T_{\infty}(p_0)}\right| 
		e^{-\alpha y_0} \dd y_0\\
	&= \bigO{1} \left(\log(k_n) n^{-(2\alpha - 1)} + k_n^2 n^{-(4\alpha - 2)} \right) 
		\int_{0}^\infty \rho(y_0,k_n) e^{-\alpha y_0} \dd y_0\\
	&= \bigO{1} \left(\log(k_n) n^{-(2\alpha - 1)} + k_n^2 n^{-(4\alpha - 2)} \right) p_{k_n} = \smallO{s(k_n) p_{k_n} k_n^2},
\end{align*}
%\TM{ Typo : $e^{-\alpha}$ }
where the last part follows from Lemma~\ref{lem:scaling_s_alpha} and the fact that $s(k_n)^2 = \smallO{s(k_n)}$. This establishes~\eqref{eq:triangle_diff_main_term} and hence finishes the proof.
%
% \TM{ Also justify penultimate step! }\PvdH{What do you mean?}
%To finish the argument we observe that
%\begin{align*}
%	\int_{a_n^-}^{a_n^+} \rho(y,k_n) P(y) e^{-\alpha} \dd y
%	&= \bigT{1} \Exp{N_{\infty}(k_n)} \gamma(k_n) = \bigT{ s(k_n) k_n^{-(2\alpha + 1)}}.
%\end{align*}
\end{proof}

From the proof of Proposition~\ref{prop:convergence_average_clustering_P_n} we obtain the following useful corollary, which will be used in Section~\ref{sec:concentration_c_P_n}.

@@

%Define
%\begin{equation}\label{eq:def_tilde_triangle_indicator}
%	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal}
%\end{equation}
%Then
%\[
%	\sum_{p_1,p_2 \in \Pcal_n}^{\ne} T_{\text{box}}(p_0,p_1,p_2) - \widetilde{T}_{\text{box}}(p_0,p_1,p_2)
%	= \sum_{p_1,p_2 \in \Pcal_n}^{\ne} \ind{p_1 \in \BallPon{p_0}} \ind{p_2 \in \mathcal{T}_{\Pcal \Delta \Pcal_n}(p_0,p_1)},
%\]
%where the sums are over all distinct pairs $(p_1, p_2) \in \Pcal_n$.

%\TM{ Note you cannot really force/assume $p_0 = (0,y)$ is in your PPP $\Pcal_n$. (The summation over $\Pcal_n \setminus (0,y)$ suggests this is what you do.) For taking the expectation of sort of thing, Palm theory / Mecke are used typically. At this point though you can again still just say we add $p_0$ to the PPP. }\PvdH{I changed the text at the beginning of this section to mention this. Since we now consider the point process $\Pcal_n \cup p_0$ the above summation is over $\Pcal_n$.}

@@

We now come back to computing the expected number of triangles attached to node at height $y$ in $\Gbox$ that are not present in $\Ginf$.

@@

Recall that $\mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$ denotes the region of points which form triangles with $p$ and $p^\prime$ in $\Gbox$ but not in $\Ginf$. Figure \ref{fig:comparing_triangles_diff_intersections} shows an example of a configuration where $\mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime) \ne \emptyset$. We observe that $\mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime) \ne \emptyset$ because the right boundary of the ball $\BallPon{p^\prime}$ exists the right boundary of the box $\Rcal$ and then, since we identified the boundaries, continues from the left so that $\BallPon{p^\prime}$ covers part of the ball $\BallPon{p}$ which would not be covered in the infinite limit model.

>>>>

Recall that $\Gbox$ is obtained by restricting the Poisson point process $\PPP$ to the box $\Rcal = (-I_n, I_n] \times (0, R_n]$, with $I_n = \frac{\pi}{2} e^{R_n/2}$ and connecting two points $p_1, p_2 \in \Rcal$ if and only if $|x_1 - x_2|_{\pi e^{R_n/2}} \le e^{(y_1 + y_2)/2}$. We also recall that by definition of the norm $|.|_{\pi e^{R_n/2}}$ the left and right boundaries of $\Rcal$ are identified. See Section~\ref{ssec:finite_model} for more details. Due to this identification of the boundaries some triples of nodes that form a triangle in the finite box model do not form a triangle in the infinite model. Therefore, to establish the required result we need to compute the asymptotic difference between triangle counts in both models.

@@

%We now turn to the task of calculating the expected number of triangles of a node at height $y$, for both the infinite model and the finite box model. 
For any $p \in \R \times \R_+$ we define for the finite box model,
\[
	T_{\text{box}}(p) = \sum^{\ne}_{p_1, p_2 \in \Pcal \setminus \{p\}} T_{\text{box}}(p,p_1,p_2)
\]
where the sum is over all distinct pairs in $\Pcal \setminus p$ and
\[
	T_{\text{box}}(p,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPon{p_1}}.
\]
Similarly, for the infinite model we define
\[
	T_\infty(y) = \sum_{p_1, p_2 \in \Pcal \setminus (0,y)}^{\ne} T_\infty(y,p_1,p_2),
\]
where
\[
	T_{\infty}(y,p_1,p_2) = \ind{p_1 \in \BallPo{y}}\ind{p_2 \in \BallPo{y}}\ind{p_2 \in \BallPo{p_1}}.
\]
Here we slightly abuse notation and write $\BallPo{y}$ for $\BallPo{(0,y)}$. We will adopt this notational convention throughout the remainder of this section, to keep notation concise.

We will first relate $\gamma(k_n)$ and $\Exp{c^\ast(k_n; \Gbox)}$ using $T_\infty(y)$ and $T_{\text{box}}(y)$. Recall the definition of $\Kcal_{C}(k_n)$
\[
	\Kcal_C(k_n) = \left\{y \in \R_+ : \frac{k_n - C \sqrt{k_n \log(k_n)}}{\xi} \le e^{\frac{y}{2}}
	\le \frac{k_n + C \sqrt{k_n \log(k_n)}}{\xi} \right\},	
\]

@@

For~\eqref{eq:alt_c_ast_kn} we recall that
\[
	c^\ast(k_n; \Gbox) = \frac{1}{\Exp{\Nbox(k_n)}} \sum_{p \in \Pcal} c_{\text{box}}(p)\ind{\Dbox(p) = k_n},
\]
where $c_{\text{box}}(p)$ can be expressed as
\[
	c_{\text{box}}(p) = \frac{1}{\binom{\Dbox(p)}{2}} \sum_{p_1, p_2 \in \Pcal \setminus p}^{\ne} T_{\text{box}}(p,p_1,p_2)
	= \frac{T_{\text{box}}(p)}{\binom{\Dbox(p)}{2}}.
\]
By the Campbell-Mecje formula
\begin{align*}
	\Exp{c^\ast(k_n; \Gbox)} 
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \Exp{c_{\text{box}}(p)\ind{\Dbox(p) = k_n}} f(x,y) \dd x \dd y\\
	&= \frac{1}{\Exp{\Nbox(k_n)}} \int_{\Rcal} \CExp{c_{\text{box}}(p)}{\Dbox(p) = k_n}\rho_{\text{box}}(p,k_n) 
		f(x,y) \dd x \dd y\\
	&= (1+\smallO{1})\frac{n}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
		\rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line follows from a concentration of heights argument, for which we used that $\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n} \le 1$. To analyze the conditional expectation we observe that, similar to the analysis of $\gamma(k_n)$, conditioned on there being $k_n$ points in $\BallPon{y}$, each point $u_i = (x_i,y_i)$ is independently distributed according to the probability density $\Mu{\BallPon{y}}^{-1} \ind{u_i \in \BallPon{y}} f(x_i,y_i)$. Therefore,
\begin{align*}
	\CExp{c_{\text{box}}(y)}{\Dbox(y) = k_n}
	&= \binom{k_n}{2}^{-1} \Exp{\sum_{1 \le i < j \le k_n} \ind{u_i \in \BallPon{u_j}}} \\
	&= \Exp{u_1 \in \BallPon{u_2}} \\
	&= \Mu{\BallPon{y}}^{-2} \iint T_{\text{box}}(y,p_1,p_2) f(x_1,y_1) f(x_2,y_2) 
		\dd x_1 \dd y_1 \dd x_2 \dd y_2\\
	&= \Mu{\BallPon{y}}^{-2} \Exp{T_{\text{box}}(y)}.
\end{align*}
and thus, by applying a concentration of heights argument on $\Mu{\BallPon{y}}^{-2}$,
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{n \Mu{\BallPon{2\log(k_n/\xi)}}^{-2}}{\Exp{\Nbox(k_n)}} \int_{\Kcal_C(k_n)} \Exp{T_{\text{box}}(y)} \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
To finish the argument, we first note that $\Mu{\BallPon{2\log(k_n/\xi)}}^{-2} = (1+\smallO{1})k_n^2$, while
\[
	\Exp{N_{\text{box}}(k_n)} = \int_{\Rcal} \rho_{\text{box}}(y,k_n) f(x,y) \dd x \dd y,
\]
so that by a concentration of heights argument,
\[
	\Exp{N_{\text{box}}(k_n)} = (1+o(1)) n \int_{0}^\infty \rho(y,k_n) \alpha e^{-\alpha y} \dd y 
	= (1+\smallO{1}) n p_{k_n}.
\]
We therefore conclude that
\[
	\Exp{c^\ast(k_n; \Gbox)} = (1 + \smallO{1}) \frac{1}{k_n^2 p_{k_n}} \int_{\Kcal_C(k_n)} 
		\Exp{T_{\text{box}}(y)} \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\]

@@

To analyze $T_{\text{box}}(y_0) - T_\infty(y_0)$ we first reiterate that the difference between the indicator $\ind{p_1 \in \BallPon{p}}$ in the finite box model and $\ind{p_1 \in \BallPo{p}}$ is that in $\Gbox$ we identified the boundaries of the interval $[-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2} e^{R_n/2}]$ and we stop at height $y = R_n$. This induces a difference in triangle counts between both models. 
To see this, note that for any $p = (x,y)$ with $0 \le y \le R_n$ we have that $\BallPon{p} = \BallPo{p} \cap \Rcal$. This means that if $p^\prime, p_2 \in \BallPon{p}$ and $p_2 \in \BallPo{p^\prime} \cap \Rcal$ then $p_2 \in \BallPon{p} \cap \BallPon{p^\prime}$ and hence $(p,p^\prime,p_2)$ form a triangle both in $\Gbox$ and $\Ginf$. However, it could happen that there are points in the intersection $\BallPon{p} \cap \BallPon{p^\prime}$ that are not in $\BallPo{p} \cap \BallPo{p^\prime}$. Let us denote this region by $\mathcal{T}(p,p^\prime)$, see Figure~\ref{fig:comparing_triangles_diff_intersections} for an example of this region. Then, any $p_2 \in \mathcal{T}(p,p^\prime)$ creates a triangle with $p$ and $p^\prime$ in $\Gbox$ that is not present in $\Ginf$. Finally, any point $p_2 \in \BallPo{p} \cap \BallPo{p^\prime}$ with height $y_2 > R$ creates a triangle with $p, p^\prime$ in $\Ginf$ but not in $\Gbox$.

Let us now define the following triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus \{p_0\}}^{\ne} 
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal}.
\]
Then $\widetilde{T}_{\text{box}}(p_0)$ only counts those triangles attached to $p_0$ that exist in both $\Gbox$ and $\Ginf$ and thus, by definition of the region $\mathcal{T}(p_0,p_1)$,
\[
	T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)
	= \sum_{p_1, p_2 \in \Pcal \setminus \{p_0\}}^{\ne} \ind{p_1 \in \BallPon{p_0}} 
			\ind{p_2 \in \mathcal{T}(p_0, p_1)}.
\]
%Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.
%Observe that the difference between $\widetilde{T}_{\text{box}}(p_0)$ and $T_{\text{box}}(p_0)$ is in the last indicator in $\widetilde{T}_{\text{box}}(p_0,p_1,p_2)$.  counts those points $p_2$ in the box $\Rcal$ that are in the neighbourhood $\BallPo{p_1}$ of $p_1$ in the infinite model. Thus, $\widetilde{T}_{\text{box}}(p_0)$ only count those triangles $(p_0,p_1,p_2)$ that also exist in $\Ginf$.

The next result, which is crucial for the proof of Proposition~\ref{prop:convergence_average_clustering_P_n}, computes the expected measure of $\mathcal{T}(p,p^\prime)$ with respect to $p^\prime$.

\begin{lemma}\label{lem:clustering_error_T_term}
Let $p_0 = (0,y)$ with $y \in \Kcal_{C}(k_n)$. Then as $n \to \infty$,
\[
	\Exp{\left|T_{\text{box}}(p_0) - \widetilde{T}_{\text{box}}(p_0)\right|}
	= y \, \bigO{n^{-(2\alpha - 1)}} + e^{y}\,\bigO{n^{-(4\alpha-2)}}.
\]
%\[
%	\int_{\Rcal} \mu\left(\mathcal{T}(p,p^\prime)(p_0,p_1)\right) f(x_1,y_1) 
%	\dd x_1 \dd y_1 = \bigO{y n^{-(2\alpha - 1)} + n^{-(2\alpha-1)} e^{y}}
%\]
\end{lemma}

@@

\begin{proof}[Proof of Proposition~\ref{prop:convergence_average_clustering_P_n}]
Let us write $\Rcal^\prime := (\R \times \R_+) \setminus \Rcal$ and let $p_0 = (0,y)$ denote the typical point. Next we recall that it is enough to show~\eqref{eq:triangle_diff_main_term}, so that in particular we have that $y \in \Kcal_{C}(k_n)$.

@@

\TM{ in the 1st integral we could have restricted to the ball of $p$, and possibly saved a lot. Does that not help? }\PvdH{the first integral is taken care of by another Lemma so it does not matter I think.}
The first part is taken care of by Lemma~\ref{lem:clustering_error_T_term}. For the other integral we have
\begin{align*}
	&\hspace{-30pt}\iint_{\Rcal^\prime} T_{\infty}(p_0,p_1,p_2) f(x_1,y_1) f(x_2,y_2)
		\dd x_2 \dd y_2 \dd x_1 \dd y_1\\
	&\le \left(\int_{\Rcal^\prime} \ind{p_1 \in \BallPo{p_0}} f(x_1,y_1) \dd x_1 \dd y_1\right)^2
		= \bigO{\left(e^{y/2} \int_{R_n}^\infty e^{-(\alpha - \frac{1}{2})y_1} \dd y_1\right)^2}\\
	&= \bigO{e^{y} e^{-(2\alpha-1)R_n}} = \bigO{e^y n^{-(4\alpha - 2)}}.
\end{align*}
%\TM{ Typo : $f_{\mu, \nu}$? Also give more details how line 3 follows. }
Thus we conclude, using Lemma~\ref{lem:clustering_error_T_term}, that,
\begin{equation}\label{eq:triangle_count_diff_scaling}
	\left|\Exp{T_{\text{box}}(p_0) - T_{\infty}(p_0)}\right| = \bigO{y n^{-(2\alpha - 1)} + n^{-(4\alpha-2)} e^{y}}.
\end{equation}
Therefore, on $\Kcal_C(k_n)$, 
\begin{align*}
	&\hspace{-30pt}\int_{\Kcal_{C}(k_n)} \rho(y_0,k_n) \left|\Exp{T_{\text{box}}(p_0) - T_{\infty}(p_0)}\right| 
		e^{-\alpha y_0} \dd y_0\\
	&= \bigO{1} \left(\log(k_n) n^{-(2\alpha - 1)} + k_n^2 n^{-(4\alpha - 2)} \right) 
		\int_{0}^\infty \rho(y_0,k_n) e^{-\alpha y_0} \dd y_0\\
	&= \bigO{1} \left(\log(k_n) n^{-(2\alpha - 1)} + k_n^2 n^{-(4\alpha - 2)} \right) p_{k_n} = \smallO{s(k_n) p_{k_n} k_n^2},
\end{align*}
%\TM{ Typo : $e^{-\alpha}$ }
where the last part follows from Lemma~\ref{lem:scaling_s_alpha} and the fact that $s(k_n)^2 = \smallO{s(k_n)}$. This establishes~\eqref{eq:triangle_diff_main_term} and hence finishes the proof.
%
% \TM{ Also justify penultimate step! }\PvdH{What do you mean?}
%To finish the argument we observe that
%\begin{align*}
%	\int_{a_n^-}^{a_n^+} \rho(y,k_n) P(y) e^{-\alpha} \dd y
%	&= \bigT{1} \Exp{N_{\infty}(k_n)} \gamma(k_n) = \bigT{ s(k_n) k_n^{-(2\alpha + 1)}}.
%\end{align*}
\end{proof}

From the proof of Proposition~\ref{prop:convergence_average_clustering_P_n} we obtain the following useful corollary, which will be used in Section~\ref{sec:concentration_c_P_n}. Recall that
\[
	\rho_{\mathrm{box}}(y,k_n) = \Prob{\Po(\mu(\BallPon{y})) = k_n},
\]
denote the degree distribution of a point $p_0 = (0,y)$ in $\Gbox$.

@@

%Define
%\begin{equation}\label{eq:def_tilde_triangle_indicator}
%	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal}
%\end{equation}
%Then
%\[
%	\sum_{p_1,p_2 \in \Pcal}^{\ne} T_{\text{box}}(p_0,p_1,p_2) - \widetilde{T}_{\text{box}}(p_0,p_1,p_2)
%	= \sum_{p_1,p_2 \in \Pcal}^{\ne} \ind{p_1 \in \BallPon{p_0}} \ind{p_2 \in \mathcal{T}(p,p^\prime)(p_0,p_1)},
%\]
%where the sums are over all distinct pairs $(p_1, p_2) \in \Pcal$.

%\TM{ Note you cannot really force/assume $p_0 = (0,y)$ is in your PPP $\Pcal$. (The summation over $\Pcal \setminus (0,y)$ suggests this is what you do.) For taking the expectation of sort of thing, Palm theory / Mecke are used typically. At this point though you can again still just say we add $p_0$ to the PPP. }\PvdH{I changed the text at the beginning of this section to mention this. Since we now consider the point process $\Pcal \cup p_0$ the above summation is over $\Pcal$.}

@@

We now come back to computing the expected number of triangles attached to a node at height $y$ in $\Gbox$ that are not present in $\Ginf$.

@@

Recall that $\mathcal{T}(p,p^\prime)$ denotes the region of points which form triangles with $p$ and $p^\prime$ in $\Gbox$ but not in $\Ginf$. Figure \ref{fig:comparing_triangles_diff_intersections} shows an example of a configuration where $\mathcal{T}(p,p^\prime) \ne \emptyset$. We observe that $\mathcal{T}(p,p^\prime) \ne \emptyset$ because the right boundary of the ball $\BallPon{p^\prime}$ exits the right boundary of the box $\Rcal$ and then, since we identified the boundaries, continues from the left so that $\BallPon{p^\prime}$ covers part of the ball $\BallPon{p}$ which would not be covered in the infinite limit model.

To further analyze this, let us introduce some notation. For any $p = (x,y) \in \Rcal$ we will define the left and right boundary functions as, respectively,
\begin{align}
	b_p^-(z) &= \begin{cases}
		2 \log\left(x-z\right) - y &\mbox{if }  -\frac{\pi}{2} e^{R_n/2} \le z \le x - e^{y/2}  \\
		2\log\left(\pi e^{R_n/2} + x - z\right) - y 
			&\mbox{if } x - e^{(y + R_n)/2} + \pi e^{R_n/2} \le z \le \frac{\pi}{2} e^{R_n/2}\\
		0 &\mbox{otherwise}
	\end{cases}\\
	b_p^+(z) &= \begin{cases}
		2 \log\left(z-x\right) - y &\mbox{if } x + e^{y/2} \le z \le \frac{\pi}{2} e^{R_n/2} \\
		2\log\left(\pi e^{R_n/2} + z - x\right) - y 
			&\mbox{if } -\frac{\pi}{2} e^{R_n/2} \le z \le x + e^{(y + R_n)/2} - \pi e^{R_n/2}\\
		0 &\mbox{otherwise}
	\end{cases}
\end{align}
Note that these functions describe the boundaries of the ball $\BallPon{p}$. In particular, $p^\prime = (x^\prime, y^\prime) \in \BallPon{p}$ if and only if $y^\prime \ge \min\left\{b_p^-(x^\prime), b_p^+(x^\prime)\right\}$.
